---
title: "Extracting dispersing events"
author: "Aurore Ponchon"
format: html
editor: visual
---

## 1- Formatting data to constitute a full long dataset

Data are stored in .xls or .xlsx files. They consist of blocks of data encompassing groups from the same region. A first formatting of the tables is made to ensure that each group name is the first line of a block. Region names are isolated before blocks so that they are automatically assigned to following blocks.

Duplicates are checked in names and region names, as well as duplicates in Individual names (same tag but different groups for the same observation file).

```{r}
#| echo: false
#| message: false
#| warning: false

library(here)
library(tidyverse)
library(readxl)
library(writexl)
library(purrr)

source(here("R","extract_functions.R"))

#identify particular status in Grupo column
dispersers<-c("OUT","Out","out","disp","Disp","DISP","OUT PT","Grace-disp
","Alex-disp","28/01/13")
solo<-c("Sozinha","Sozinho","sozinha","sozinho","SOZINHA","soz.","SOZINHO",
        "Expulsa","Soz.")
dead<-c("DEAD","died","Died","DIED")
names<-c("chiquinho","Georgia","Luisa","Antônia","saiu","Luana",
         "Camilo","André")
#list all files
files<-list.files(here("data","RawData","Groups"),
                  pattern = "xls",recursive=T,
                  full.names = T)

#function to read each file separately

# DATA<-data.frame()
# 
# for (i in 1:length(files)){
#   
#   print(i)
#   
#   xx<-extract_long_data(files[i])
#   
#   DATA<-rbind(DATA,xx)
# }

DATA<- purrr::map_dfr(files, extract_long_data )

reg<-data.frame(Region=sort(unique(DATA$Region)))
group<-data.frame(Group=sort(unique(DATA$Group)))
  

#write_xlsx(DATA, path=here("Data","NewlyCreatedData","DAT.xlsx"))

save(DATA,file=here("Data","NewlyCreatedData","all_long.RData"))


subseto<-DATA %>% 
  filter(Group=="DN"|
         Group=="Fugitivo"|
         Group=="Fusão"|
         Group=="GM"|
           Group=="FQL"|
           Group=="Jaguar"|
           Group=="ML2"|
           Group=="MR"|
           Group=="NA") %>% 
  select(-Grupo)

write_xlsx(subseto,path=here("Data","NewlyCreatedData","check_group_Valeria.xlsx"))
```

# Selecting individuals to include in observation dataset

The dataset for capture-recapture analyses includes individuals which have been clearly identified. So dispersed individuals and dead individuals are removed.

```{r}

load(here("Data","NewlyCreatedData","all_long.RData"))


dat<-DATA %>% 
  dplyr::filter(Disp=="0" & Death=="0") %>% 
  dplyr::filter(GLT!="?" & GLT !="?-3" & GLT!="IN" & GLT!="FT" &
                  GLT!="T0") %>% 
  dplyr::arrange(DateObs) %>% 
  dplyr::select(DateObs,Region,Group, GLT, Sexo ) %>% 
  dplyr::distinct() %>% 
  pivot_wider(names_from=DateObs,
              names_prefix="Obs",
              values_from=Region)


DATAT %>%
    dplyr::group_by(DateObs,GLT) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::filter(n > 1L & GLT!="?" & GLT!="T0" & GLT!="IN" & 
                    GLT!="FT")

```
