---
title: "Extracting dispersing events"
author: "Aurore Ponchon"
format: html
editor: visual
---

## 1- Formatting data to constitute a full long dataset

Data are stored in .xls or .xlsx files. They consist of blocks of data encompassing groups from the same region. A first formatting of the tables is made directly within the excel files to ensure that each group name is the first line of a block. Region names are isolated before blocks so that they are automatically assigned to following blocks.

```{r}
#| echo: false
#| message: false
#| warning: false

library(here)
library(tidyverse)
library(readxl)
library(xlsx)
library(conflicted)
library(progressr)
conflict_prefer("mutate", "dplyr")


##Get fun progress bars in scripts
handlers("progress")
red_heart <- cli::col_green(cli::symbol$tick)
handlers(handler_txtprogressbar(char = red_heart))



source(here::here("R","extract_functions.R"))
source(here::here("R","revalue_variables.R"))

#identify particular status in Grupo column
dispersers<-c("OUT","Out","out","disp","Disp","DISP","OUT PT","Grace-disp
","Alex-disp","28/01/13", "saiu","Expulsa")
solo<-c("Sozinha","Sozinho","sozinha","sozinho","SOZINHA","soz.","SOZINHO",
        "Soz.")
dead<-c("DEAD","died","Died","DIED")
names<-c("chiquinho","Georgia","Luisa","Antônia","Luana",
         "Camilo","André")
#list all files
files<-list.files(here::here("data","RawData","Groups"),
                  pattern = "xls",recursive=T,
                  full.names = T)

#function to read each file separately

# DATA<-data.frame()
# 
# for (i in 1:length(files)){
#   
#   print(i)
#   
#   xx<-extract_long_data(files[i])
#   
#   DATA<-rbind(DATA,xx)
# }

DATA<- purrr::map_dfr(files, extract_long_data,.progress=T)

data.clean<-DATA %>% 
            revalue_region() %>% 
      #      revalue_birth() %>%    #issue with FA20 (2 birth dates!!)
            revalue_stage() %>% 
            revalue_sex() %>% 
            revalue_name() %>% 
            dplyr::arrange(GLT,DateObs)


sort(unique(data.clean$Region))
sort(unique(data.clean$Group))
  
data.clean %>% 
  
    # dplyr::filter(GLT!="?" & 
    #                 GLT !="?-3" & 
    #                 GLT!="IN" & 
    #                 GLT!="FT" &
    #                 GLT!="T0" & 
    #                 Group!="?") %>% 
   # dplyr::select(DateObs,Region, Group, GLT, Sexo,Idade,Disp,Death,Solo) %>% 
    write.table(., file=here::here("Data","NewlyCreatedData","DATA.csv"),
               quote=F,row.names=F ,sep=";")

save(data.clean,file=here::here("Data","NewlyCreatedData","all_long.RData"))


# subseto<-DATA %>% 
#   filter(Group=="DN"|
#          Group=="Fugitivo"|
#          Group=="Fusão"|
#          Group=="GM"|
#            Group=="FQL"|
#            Group=="Jaguar"|
#            Group=="ML2"|
#            Group=="MR"|
#            Group=="NA") %>% 
#   select(-Grupo)
# 
# write_xlsx(subseto,
# path=here("Data","NewlyCreatedData","check_group_Valeria.xlsx"))
```

# Checking duplicates and errors

Duplicates are checked in group and region names, as well as in individuals. Individual variables are also filled, to ensure that they are consistent among observations of the same individual (notably sex and date birth). After data corection, all objects should be empty!

```{r}
#| echo: false
#| message: false
#| warning: false

load(here::here("Data","NewlyCreatedData","all_long.RData"))


bad<-c("?","T0","IN","AF?","?-3","FT")

#identify duplicates in individual names
data.clean %>%
    #dplyr::filter(Disp=="0" & Death=="0") %>% 
    dplyr::group_by(DateObs,GLT) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::filter(n > 1L & GLT!="?" & GLT!="T0" & GLT!="IN" & 
                    GLT!="FT")

dup<-data.clean %>%
    dplyr::filter(Disp=="0" & Death=="0" & GLT!="?" & GLT!="T0" & GLT!="IN" & 
                    GLT!="FT") %>% 
    dplyr::group_by(DateObs,GLT) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::filter(n > 1L) # & GLT!="?" & GLT!="T0" & GLT!="IN") %>% 

dups<-data.clean %>% 
  dplyr::filter(GLT %in% dup$GLT) %>% 
  dplyr::arrange(GLT, DateObs)

#1377 ok
#FT = weird!

# write_xlsx(dups,
#  path=here::here("Data","NewlyCreatedData","duplicated_individuals.xlsx"))


##checking duplicates in Sex
sexo<-data.clean %>%
    dplyr::group_by(GLT) %>%
    dplyr::summarise(Sexy = Sex,.groups="keep") %>% 
    distinct() %>% 
  dplyr::filter(n()>1)
   
#MP475 problematic (same birth and both M & F)
#1375 with 3 different dates!!
#AF22, AF23, AF24, AF25 born after observation and dispersed?!
#check FA40 (2018 & 2019)
#check FT (Part of duplicate names)
#JP23 duplicate
#check 1311, 1312, MP104, MP108, MP109, MP176, MP292, MP313, MP99, TM20, UR4


#check differences between GLT and tattoo
ind<-data.clean %>% 
  dplyr::filter(!is.na(Tattoo)) %>% 
  dplyr::filter(GLT!=Tattoo) %>% 
  dplyr::arrange(GLT,DateObs)


```

# Selecting individuals to include in observation dataset

The dataset for capture-recapture analyses includes individuals which have been clearly identified. Dispersed individuals and dead individuals are removed, as well as non identified infants.

Duplicates are checked to make sure each GLT and tattoo correspond to only one unique individual. Duplicate individual tattoos are renamed in the code, not in the files.

```{r}

load(here::here("Data","NewlyCreatedData","all_long.RData"))



## Produce a table for CMR data
dat<-data.clean %>% 
  dplyr::filter(Disp=="0" & Death=="0") %>% 
  dplyr::filter(GLT!="?" & GLT !="?-3" & GLT!="IN" & GLT!="FT" &
                  GLT!="T0" & Group!="?") %>% 
  dplyr::arrange(DateObs) %>% 
  dplyr::select(DateObs,Region,Group, GLT, Sexo ) %>% 
  dplyr::distinct() %>% 
  pivot_wider(names_from=DateObs,
              names_prefix="Obs",
              values_from=Region)

```
